Towards Detecting Stalkerware Through Network Traces

### 2019-03-22 paper/main.tex
Update main.tex

What causes most stalkerware to send packets? Our captures and testing show 
that there are three main categories of updating tracker state: manual, timed, 
and on tracked feature. Most free apps fall into the manual update method, 
where there is a button on a website or on a phone app that will send a signal 
packet to the stalkerware app that initializes the upload. Most paid apps also 
include a timed auto-update feature, where an update will be sent every few 
minutes. A very small subset of both advertised that it receives a "real time 
notification" every time email, texts, calls, or GPS, was changed or added to. 
Our results show that none of the stalkerware apps were able to send in real 
time, all the apps advertised as real time simply updated at a more frequent 
frequency than other apps.
\subsection{Network Signatures}
\label{network_signatures}

@@

All of the above signatures are able to be filtered out through a Wireshark filter or made into a rule for a Network Intrusion Detection System such as Snort. There are many other parts of the packets that we could attempt to match signatures for, we will explain why we cannot add more to this list.
\subsection{Traffic Results}

@@

\section{Actual Traffic Testing}

@@

>>>>

What causes most stalkerware to send packets? Our captures and testing show 
that there are three main categories of updating tracker state: manual, timed, 
and on tracked feature. Most free apps fall into the manual update method, 
where there is a button on a website or on a phone app that will send a signal 
packet to the stalkerware app that initializes the upload. Most paid apps also 
include a timed auto-update feature, where an update will be sent every few 
minutes. A very small subset of both advertised that it receives a "real time 
notification" every time email, texts, calls, or GPS, was changed or added to. 
Our results show that none of the stalkerware apps were able to send in real 
time, all the apps advertised as real time simply updated at a more frequent 
frequency than other apps.

Where do these stalkerware apps send their traffic? In general, all of our apps tested connected to a third party server that aggregated the information for display in either a website or a similar app for a separate phone. We had anticipated that at least one app would send tracking data directly to the abuser's phone, however this was not the case in testing. Perhaps setting up such a connection would be too difficult for someone who initiates intimate partner violence, or simply because it is easier to collect phone data in a dedicated server and use it for other purposes as well. Either way, all of the apps we found connected to a specific domain that was not the abuser's phone. This was both good and bad news. While it is much easier to find network signatures for packets being sent to a single destination with its own domain name, it also means that we will not be able to find an abuser through our method. This would have been an interesting tangent, unfortunately it does not work.
\subsection{Network Signatures}
\label{network_signatures}

@@

All of the above signatures are able to be filtered out through a Wireshark filter or made into a rule for a Network Intrusion Detection System such as Snort. There are many other parts of the packets that we could attempt to match signatures for, and we will explain why we cannot add more to this list in a later section.
\subsection{Traffic Results}

@@

\section{Real world Traffic Testing}

@@

Using our specific signatures, our filters did not catch any packet that could be used as stalkerware. This is to be expected as we are very sure that there are no stalkerware apps installed on our personal phones and the filters used are quite specific to the apps that we looked at. Again, these filters are very specific with their requirements and does not scale when we add more apps to our study. However, we can confirm that stalkerware traffic is unique, and we have a filter that very accurately catches stalkerware.
